Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-04-30 index.qmd
start of bede's comment

title: PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.

@@

  Coastal areas support seagrass meadows, which offer crucial ecosystem services including erosion control and carbon sequestration. However, these areas are increasingly impacted by human activities, leading to seagrass decline and habitat fragmentation. In situ surveys, traditionally performed to monitor these ecosystems face limitations on temporal and spatial coverage, particularly in intertidal zones, prompting the use of satellite data within monitoring programs. Yet, satellite remote sensing struggles with spatial and spectral resolution, making it difficult to discriminate seagrass from other macrophytes in highly heterogenous meadows. Drone images at a very high spatial resolution offer a promising solution to address challenges related to spatial heterogeneity and intrapixel mixture. This study focuses on using drone acquisitions with a ten spectral band sensor mirroring those of Sentinel-2, for mapping intertidal macrophytes and effectively discriminating between seagrass and green macroalgae. Nine drone flights were conducted at two different altitudes (12m and 120m) across heterogeneous intertidal European habitats in France and Portugal. Low altitude flights were used to train a Deep Learning classifier based on Neural Networks to discrimintate among five taxonomic classes of intertidal vegetation: Magnoliopsida (Seagrass), Chlorophyceae (Green macroalgae), Phaeophyceae (Brown algae), Rhodophyceae (Red macroalgae) and benthic Bacillariophyceae (Diatoms). Classification of drone imagery resulted in an overall accuracy of 94% across all the sites and images, covering a total area of 467 000 m². The model exhibited an accuracy of 96.4% in identifying seagrass. Importantly, seagrass and green algae can be discriminated, although they share the same pigment composition. As the algorithm was developed for a multispectral camera with ten spectral bands in the visible and near-infrared, it could be adapted to the Multi-Spectral Instrument (MSI) onboard Sentinel-2 thus offering promising perspectives for satellite remote sensing of intertidal biodiversity over lager scales. 

@@

Coastal areas are vital hotspots for marine biodiversity, with intertidal seagrass meadows playing a crucial role at the interface between the land and oceans [@unsworth2022]. Seagrass meadows provide a myriad of ecosystem services to humanity, including carbon sequestration, oxygen production, protection against sea-level rise and coastline erosion, and limitation of eutrophication. They serve as vital habitats for a diverse array of marine and terrestrial species, providing living, breeding, and feeding grounds [@gardner2018 ; @Zoffoli2022 ; @jankowska2019]. Due to the concentration of human activities in coastal zones, seagrass meadows are directly exposed to and impacted by anthropogenic pressures. Global regression and fragmentation are currently observed due to climate change, diseases, urbanization, land reclamation, dredging, competition with alien species, and reduction in water quality [@nguyen2021 ; @soissons2018 ; @orth2006 ; @lin2018 ; @duffy2019 ; @rasheed2011long ; @chefaoui2018dramatic]. Both habitat fragmentation and reduction, in turn, can severely copromise the effectiveness of ecosystem services provided by seagrass meadows. While improvements in water quality have been recently reported in Europe, allowing an overall recovery of seagrass ecosystems at local scale, many coastal waters worldwide are still subjected to strong eutrophication processes [@deSantos2019 ; @Zoffoli2021]. Coastal eutrophication has been ass ociated to excessive accumulation of green macroalgae, so-called green tides (@devlin2023nutrients). Green tides produce shade and suffocation over seagrass individuals, thus threatening the health of seagrass ecosystems [@wang2022].

The importance of seagrass meadows and the variety of ecosystem services they provide have led to the enhancement of both global and regional programs to monitor Essential Oceanic Variable (EOVs) such as seagrass composition [@Miloslavich2018], as well as Essential Biodiversity Variable (EBVs) such as seagrass taxonomic diversity, species distribution, population abundance, and phenology [@Pereira2013]. Traditionally, indicators of seagrass status have been quantified using *in situ* measurements. The acquisition of field data in intertidal zones is however notoriously challenging. Intertidal seagrass meadows are only partially exposed during low tide and can be situated in difficult-to-reach mudflats, potentially leading to inaccurate and limited estimations with conventional sampling techniques [@nijland2019]. Satellite observations have been proven effective in complementing in situ sampling, allowing for the near real-time and consistent retrieval of seagrass EOVs and EBVs over extensive meadows. [@Zoffoli2021 ; @xu2021 ; @Traganos2018 ; @coffer2023]

While satellite remote sensing (RS) provides temporally consistent observations over large spatial scales, its utilization over intertidal areas is limited by several constraints. Satellite missions with a high temporal resolution (e.g. daily MODIS observation) are limited by a too coarse spatial resolution (\>100m) to accurately map patchy seagrass meadows. Missions with a high spatial resolution such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral resolution. The limited number of spectral bands is indeed a challenge to accurately discriminate seagrass from others co-existing macrophytes. In particular Chlorophyceae (green algae) and marine Magnoliopsida (seagrass) share the same pigment composition [@ralph2002 ; @Douay2022]. Therefore, to someone not specialized in the field, their spectral signatures may appear to be alike [@Davies2023 ; @bannari2022]. Recently, using advanced machine-learning algorithms trained with a large hyperspectral library of more than 300 field reflectance spectra, @Davies2023 demonstrated that it was possible to discriminate Magnoliopsida from Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral resolution. However the application of this approach to satellite RS remains to be validated. Moreover patches of green algae can develop at small spatial scales that are not observable using non-commercial satellite imagery [@tuya2013], especially during the initial stage of a green tide.

Drones can potentially fill the gaps left by satellite RS and in situ measurements, due to their ability to provide spatially-explicit observations at very high spatial resolutions (pixel size from mm to cm) while capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping species distribution [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, most case studies were previously limited to a single flight, restricting the generalizability of their application over wider geographical scales [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. The present study aimed at analyzing the potential of multispectral drone RS to map intertidal macrophytes over a diverse biogeographical range, with a particular focus on discriminating Magnoliopsida and Chlorophyceae (Seagrass and Green Algae respectively). Ten drone flights were performed over soft-bottom intertidal areas along two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green, or red macroalgae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected for their relatively extensive intertidal seagrass beds. Two sites are located in the Gulf of Morbihan (@fig-map A), France (47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only connected to the sea through a 900m wide channel. A total of 53 small islands are scattered across the gulf leading to 250 km of shorelines. Patchy seagrass meadows can be found on a lot of these islands. Within the Gulf of Morbihan we chose two sites : one located on one of the gulf's islands (Arz island) and the other located in the southern part of the gulf (Duer). Two other sites were located in Bourgneuf Bay, France (46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed macrotidal bay, protected from waves by Noirmoutier Island. Bourgneuf bay hosts a large intertidal seagrass meadow of about 6 km². Within this meadow, the sites observed by drones (L'Epine and Barbatre, @fig-map B) contained monospecific beds of *Nanozostera noltei* (dwarf eelgrass) with very little mixing with other macrophytes. Three sites were surveyed in the Ria de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this lagoon is about 80 km² with many narrow channels, large salt marshes and many mudflats that uncovers at low tide. It is connected to the open sea through a single channel, with a tidal lag between the North and the South of the lagoon. The southernmost site (Gafanha) was a mudflat located in a tributary river of the lagoon whereas the two other sites (Mataducos and Marinha Lanzarote) were situated in the middle of the lagoon and only accessible by boat (@fig-map C). These Portuguese sites are characterized by a more diverse intertidal vegetation, where patches of seagrass intermingle with red, brown and green macroalgae.

@@

```{r Table1}
#| echo: false
#| warning: false
#| label: tbl-flights
#| tbl-cap: List of drone flight, summarising the date, the altitude and the purpose of each flight. 12 m  and 120 m flights have a spatial resolution of 8 and 80 mm respectively.
#| cache: true

@@

Before each flight, targets used as Ground Control Points (GCPs) were distributed over the study site and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). GCPs were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was also used to georeference quadrats of 0.25 m² assessing the presence or absence of five key taxonomic classes of intertidal vegetation : Bacillariophyceae (Unicellular benthic diatoms forming biofilms at the sediment surface during low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf seagrass), Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae) (@fig-vegetation). Pictures of each quadrat were uploaded online to the Global Biodiversity Information Facility (GBIF) platform [@BedeGbif], a public open portal to store and share biodiversity data. Each photograph was also processed to estimate the percent cover of each type of vegetation using an image processing software (imageJ). For all quadrats, the hyperspectral reflectance signatures of each vegetation class was measured using an ASD FieldSpec HandHeld 2 spectroradiometer, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures from quadrats serve for dual purposes: they validate the radiometric calibration of drone data and contribute to error reduction in photo interpretation. The workflow developed in this study is presented in @fig-workflow.

@@

A structure-from-motion photogrammetry software (Agisoft Metashape) was used to process images to obtain multispectral orthomosaics of each flight. The workflow for orthomosaicking was the same for every flight. First, tying key points were detected inside of each image and between overlapping images in order to obtain a sparse point cloud. This cloud was cleaned using reprojection accuracy metric in order to remove noisy points. A dense point cloud was then produced using a structure from motion algorithm. A surface interpolation of this dense point cloud was made to obtain a digital surface model (DSM), used to reconstruct the multispectral ortho-image. Low altitude drone flights produced ortho-images with a very high spatial resolution (8 mm per pixel), making it efficient to visually distinguish between the various types of vegetation. High altitude fights on the other hand allowed to cover large areas and produced images with a pixel size of 80 mm (@tbl-flights).

@@

```{r Figure3}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: "Schematic representation of the workflow. Diamonds represent input or output data, and rectangles represent Python processing algorithms. The overall workflow of this study is divided into two distinct parts based on the spatial resolution of the drone flights: high-resolution flights (pixel size: 8 mm) were utilized for training and prediction of the Neural Network model, whereas lower-resolution flights (pixel size: 80 mm) were solely employed for prediction purposes. Validation has been performed on both high and low resolution flights."
#| label: fig-workflow
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

### Neural Network model building

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Neural Network model with 2 hidden layers. The training pixels were categorized into seven different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (@tbl-flights) were used for training because their 8 mm spatial resolution allowied to avoid spectral sub-pixel mixing and to accurately identify vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights were used to train the model (@tbl-validationPX). Twenty one variables were used by the model as predictors: the ten raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same ten spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the Normalized difference vegetation index (NDVI, @eq-ndvi). Standardisation of spectral bands is commonly used to eliminate the scaling differences between spectra and to limit the effect of biomass on the spectra shape [@Douay2022 ; @Davies2023].

@@

Variable Importance Plots (VIP) serve as a method to identify which predictors are important for predicting a specific class. This is achieved by repeatedly predicting the same dataset while shuffling one predictor at a time. The benchmark score obtained after each iteration is then compared to the benchmark score obtained without shuffling any variables. The greater the difference between these two benchmark values, the more important the variable is for the model [@WEI2015399].

>>>>

title: Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

@@

  Coastal areas support seagrass meadows, which offer crucial ecosystem services including erosion control and carbon sequestration. However, these areas are increasingly impacted by human activities, leading to habitat fragmentation and seagrass decline. In situ surveys, traditionally performed to monitor these ecosystems face limitations on temporal and spatial coverage, particularly in intertidal zones, prompting the addition of satellite data within monitoring programs. Yet, satellite remote sensing struggles with spatial and spectral resolution, making it difficult to discriminate seagrass from other macrophytes in highly heterogenous meadows. Drone images at a very high spatial resolution offer a promising solution to address challenges related to spatial heterogeneity and intrapixel mixture. This study focuses on using drone acquisitions with a ten spectral band sensor mirroring those of Sentinel-2, for mapping intertidal macrophytes and effectively discriminating between seagrass and green macroalgae. Nine drone flights were conducted at two different altitudes (12m and 120m) across heterogeneous intertidal European habitats in France and Portugal. Low altitude flights were used to train a Deep Learning classifier based on Neural Networks to discrimintate among five taxonomic classes of intertidal vegetation: Magnoliopsida (Seagrass), Chlorophyceae (Green macroalgae), Phaeophyceae (Brown algae), Rhodophyceae (Red macroalgae) and benthic Bacillariophyceae (Diatoms). Classification of drone imagery resulted in an overall accuracy of 94% across all the sites and images, covering a total area of 467 000 m². The model exhibited an accuracy of 96.4% in identifying seagrass. Importantly, seagrass and green algae can be discriminated, although they share the same pigment composition. As the algorithm was developed for a multispectral camera with ten spectral bands in the visible and near-infrared, it could be adapted to the Multi-Spectral Instrument (MSI) onboard Sentinel-2 thus offering promising perspectives for satellite remote sensing of intertidal biodiversity over lager scales. 

@@

Coastal areas are vital hotspots for marine biodiversity, with intertidal seagrass meadows playing a crucial role at the interface between the land and oceans [@unsworth2022]. Seagrass meadows provide a myriad of ecosystem services to humanity, including carbon sequestration, oxygen production, protection against sea-level rise and coastline erosion, and limitation of eutrophication. They serve as vital habitats for a diverse array of marine and terrestrial species, providing living, breeding, and feeding grounds [@gardner2018 ; @Zoffoli2022 ; @jankowska2019]. Due to the concentration of human activities in coastal zones, seagrass meadows are directly exposed to and impacted by anthropogenic pressures. Global regression and fragmentation of seagrass meadows are currently observed due to climate change, diseases, urbanization, land reclamation, dredging, competition with alien species, and reduction in water quality [@nguyen2021 ; @soissons2018 ; @orth2006 ; @lin2018 ; @duffy2019 ; @rasheed2011long ; @chefaoui2018dramatic]. Both habitat fragmentation and reduction, in turn, can severely compromise the effectiveness of ecosystem services provided by seagrass meadows. While improvements in water quality have been recently reported in Europe, allowing an overall recovery of seagrass ecosystems at local scales, many coastal waters worldwide are still subjected to strong eutrophication processes [@deSantos2019 ; @Zoffoli2021]. Coastal eutrophication has been associated to excessive accumulation of green macroalgae, so-called green tides [@devlin2023nutrients]. Green tides produce shade and suffocation over seagrass individuals, thus threatening the health of seagrass ecosystems [@wang2022].

The importance of seagrass meadows and the variety of ecosystem services they provide have led to the enhancement of both global and regional programs to monitor Essential Oceanic Variable (EOVs) such as seagrass composition [@Miloslavich2018], as well as Essential Biodiversity Variable (EBVs) such as seagrass taxonomic diversity, species distribution, population abundance, and phenology [@Pereira2013]. Traditionally, indicators of seagrass status have been quantified using *in situ* measurements. However, the acquisition of these measurements in intertidal zones is notoriously challenging. Intertidal seagrass meadows are only partially exposed during low tide and can be situated in difficult-to-reach mudflats, potentially leading to inaccurate and limited estimations with conventional sampling techniques [@nijland2019]. Satellite observations have been proven effective in complementing *in situ* sampling, allowing for the near real-time and consistent retrieval of seagrass EOVs and EBVs over extensive meadows. [@Zoffoli2021 ; @xu2021 ; @Traganos2018 ; @coffer2023]

While satellite remote sensing (RS) provides temporally consistent observations over large spatial scales, its utilization over intertidal areas is limited by several constraints. Satellite missions with a high temporal resolution (e.g. daily MODIS observation) are limited by a too coarse spatial resolution (\>100m) to accurately map patchy seagrass meadows. Missions with a high spatial resolution such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral resolution. The limited number of spectral bands challenges accurate discrimination of seagrass from other co-existing macrophytes. In particular Chlorophyceae (green algae) and marine Magnoliopsida (seagrass) share the same pigment composition [@ralph2002 ; @Douay2022]. Therefore, to someone not specialized in the field, their spectral signatures may appear to be alike [@Davies2023 ; @bannari2022]. Recently, using advanced machine-learning algorithms trained with a large hyperspectral library of more than 300 field reflectance spectra, @Davies2023 demonstrated that it was possible to discriminate Magnoliopsida from Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral resolution. However the application of this approach to satellite RS remains to be validated. Moreover patches of green algae can develop at small spatial scales that are not observable using non-commercial satellite imagery [@tuya2013], especially during the initial stage of a green tide.

Drones can potentially fill the data gaps left by satellite RS and *in situ* measurements, due to their ability to provide spatially-explicit observations at very high spatial resolutions (pixel size from mm to cm) while capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping species distribution [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, most case studies were previously limited to a single flight, restricting the generalizability of their application over wider geographical scales [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. The present study aimed at analyzing the potential of multispectral drone RS to map intertidal macrophytes over a diverse biogeographical range, with a particular focus on discriminating Magnoliopsida and Chlorophyceae (Seagrass and Green Algae respectively). Ten drone flights were performed over soft-bottom intertidal areas along the Atlantic coastlines of two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green, or red macroalgae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected for their relatively extensive intertidal seagrass beds. Two sites were located in the Gulf of Morbihan (@fig-map A), France (47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only connected to the sea through a 900m wide channel. A total of 53 small islands are scattered across the gulf leading to 250 km of shorelines. Patchy seagrass meadows can be found on many of these islands. One of the sites within the gulf was on one its islands (Arz) and the other was located further south on a mainland beach area (Duer). Two other sites were located in Bourgneuf Bay, France (46.9849°N, 2.1488°W). This is a 340 km² semi-enclosed macrotidal bay, protected from waves by Noirmoutier Island. Bourgneuf bay hosts a large intertidal seagrass meadow of about 6 km² [@ZOFFOLI2020112020]. Within this meadow, the sites observed by drones (L'Epine and Barbatre, @fig-map B) contained monospecific beds of *Nanozostera noltei* (dwarf eelgrass) with very little mixing with other macrophytes. Three sites were surveyed in the Ria de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this lagoon is \~80 km² with many narrow channels, large salt marshes and many mudflats that uncover at low tide. It is connected to the open sea through a single channel, with a tidal lag between the North and the South of the lagoon. The southernmost site (Gafanha) was a mudflat located in a tributary river of the lagoon whereas the two other sites (Mataducos and Marinha Lanzarote) were situated in the middle of the lagoon and only accessible by boat (@fig-map C). These Portuguese sites are characterized by a more diverse intertidal vegetation, where patches of seagrass intermingle with red, brown, and green macroalgae.

@@

```{r Table1}
#| echo: false
#| warning: false
#| label: tbl-flights
#| tbl-cap: List of drone flight, summarising the date, the altitude, and the purpose of each flight. 12 m  and 120 m flights have a spatial resolution of 8 and 80 mm respectively.
#| cache: true

@@

Before each flight, targets used as ground control points were distributed over the study site and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). Ground control points were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. A dGPS was also used to georeference quadrats of 0.25 m² ,which assessed the presence or absence of five key taxonomic classes of intertidal vegetation : Bacillariophyceae (Unicellular benthic diatoms forming biofilms at the sediment surface during low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf seagrass), Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae) (@fig-vegetation). Pictures of each quadrat were uploaded online to the Global Biodiversity Information Facility (GBIF) platform [@BedeGbif], a public open portal to store and share biodiversity data. Each photograph was also processed to estimate the percent cover of each type of vegetation using an image processing software [@schneider2012nih, ImageJ]. Hyperspectral reflectance signatures of each vegetation class were recorded using an ASD FieldSpec HandHeld 2 spectroradiometer, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures served dual purposes: they validate the radiometric calibration of drone data and contribute to error reduction in photo interpretation.

@@

A structure-from-motion photogrammetry software (Agisoft Metashape) was used to process images to obtain multispectral orthomosaics of each flight. The process for orthomosaicking was identical for every flight. First, tying key points were detected inside of each image and between overlapping images in order to obtain a sparse point cloud. This cloud was cleaned using reprojection accuracy metric in order to remove noisy points. A dense point cloud was then produced using a structure from motion algorithm. A surface interpolation of this dense point cloud was made to obtain a digital surface model (DSM), used to reconstruct the multispectral ortho-image [@nebel2020review]. Low altitude drone flights produced ortho-images with a very high spatial resolution (8 mm per pixel), making it efficient to visually distinguish between the various types of vegetation. High altitude fights on the other hand allowed to cover large areas and produced images with a pixel size of 80 mm (@tbl-flights).

@@

```{r Figure3}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: "Schematic representation of the workflow. Parallelograms represent input or output data, and rectangles represent Python processing algorithms. The overall workflow of this study is divided into two distinct parts based on the spatial resolution of the drone flights: high-resolution flights (pixel size: 8 mm) were utilized for training and prediction of the Neural Network model, whereas lower-resolution flights (pixel size: 80 mm) were solely employed for prediction purposes. Validation has been performed on both high and low resolution flights."
#| label: fig-workflow
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

### Training dataset building

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Neural Network model. The training pixels were categorized into seven different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (@tbl-flights) were used for training because their 8 mm spatial resolution allowied to avoid spectral sub-pixel mixing and to accurately identify vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights were used to train the model (@tbl-validationPX). Twenty one variables were used by the model as predictors: the ten raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same ten spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the Normalized difference vegetation index (NDVI, @eq-ndvi). Standardisation of spectral bands is commonly used to eliminate the scaling differences between spectra and to limit the effect of biomass on the spectra shape [@Douay2022 ; @Davies2023].

@@

Variable Importance Plots (VIP) serve as a method to identify which predictors are important for predicting a specific class. Out of the twenty one predictors utilized in this study, Variable Importance in Projection (VIP) was computed only for the raw and standardized values of the 10 spectral bands captured by the MicaSense camera. This is achieved by repeatedly predicting the same dataset while shuffling one predictor at a time. The benchmark score obtained after each iteration is then compared to the benchmark score obtained without shuffling any variables. The greater the difference between these two benchmark values, the more important the variable is for the model [@WEI2015399].
